SafeNet: A European database of plant-pollinator interactions

### 2024-03-04 Manuscript/Ms.qmd
phylo plants updated

FIGURE 3 POLLINATOR COVERAGE

```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 3}. a) Phylogenetic tree of plant families in the Safenet database, showing the number of species per family in logarithmic scale (blue circles) and the total number of interactions recorded per family (gray bars). b) Comparison of the top 50 families with the highest number of species in the SafeNet database against the total number of species recorded for those families at the European level. Species counts are displayed in logarithmic scale for better visualization.", fig.width=8, fig.height=8}

@@

#Save data to plot it in markdown file
c = readRDS("../Data/Working_files/Figures/data_figure3b.rds")
tree_family_interactions = readRDS("../Data/Working_files/Figures/data_figure3a1.rds")
plant_fam_interactions = readRDS("../Data/Working_files/Figures/data_figure3a2.rds")

# The circular layout tree
p = ggtree(tree_family_interactions, 
    layout="fan", size=0.1, open.angle=5, alpha=0.5)+ 
geom_tippoint(aes(size= Family_species), colour='cyan4') +
geom_tiplab(linetype='dashed', linesize=.05, 
    size=1.75, color= "black", offset = -45, fontface="bold") +
labs("log(Species)")+
scale_size(name = "log(Species #)") +
ggtitle("a)") +
theme(plot.title = element_text(vjust = -31, hjust = 0.12, size = 14),
      legend.position = "bottom")

#Create barplot around circular tree
p1 = p + 
new_scale_fill() +
geom_fruit(data=plant_fam_interactions, geom=geom_bar,
   mapping=aes(y=Plant_family1, x=Family_interactions1),
   pwidth=0.38, 
   orientation="y", 
   stat="identity", 
   color = "black") +
theme(legend.margin=margin(0,0,0,0),
   legend.box.margin=margin(-50,-50,-50,-50),
   plot.margin = margin(-15, -15, -10, -10, "pt")) +
scale_size(name = "log(species number)", range=c(0,4))

p2 = ggplot() + 
geom_col(data = c, 
  aes(x = reorder(Family_names, Spp_per_family), 
  y = log(Spp_per_family), fill = Group), position=position_dodge(0.1),width=1) +
xlab(NULL) +
ylab("log(species number)") +
theme_bw() +
theme(axis.text.y = element_text(size = 5, face = "bold", color = "black"),
      panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
coord_flip(expand = F) +
scale_fill_manual(name="", values = c("black", "cyan4")) +
  theme(plot.margin = margin(-10, 0, -10, -10, "pt")) +
ggtitle("b)")+
theme(legend.margin=margin(0,0,0,0),
   legend.box.margin=margin(-10,-10,-10,-10)) +
  theme(legend.key.size = unit(0.3, 'cm'),
        legend.position = "bottom",
        plot.title = element_text(size=14))

p2_1 = plot_grid(NULL,p2,NULL, ncol = 1, rel_heights  = c(1, 1.5, 0.5),rel_widths  = c(0.1, 2, 0.1))
plot_grid(p1,p2_1, rel_widths = c(16, 7), rel_heights  = c(9, 1), align = "none")

>>>>

```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.width=10, fig.height=10, fig.cap= "\\textbf{Figure 2}. Phylogenetic representation of bee genera in the SafeNet database within the context of all European bee genera. The number of interactions recorded per genus in the database is illustrated using circles, with their sizes proportional to the number of interactions on a logarithmic scale. Additionally, a gradient of colors ranging from yellow to dark purple aids in this visualization. The coverage of species recorded in SafeNet per genus at the European level is depicted with orange and light grey bars, representing the percentage of species included and not included in the database, respectively, out of the total number of bee species per genus at the European level. Grey bars indicate the total number of species per genus at the European level."}

#Load libraries
library(ggplot2)
library(ggnewscale)
library(ggtreeExtra)
library(ggtree)
library(ggpattern)
library(patchwork)
library(cowplot)

#Read data
tree_family_interactions= readRDS("../Data/Manuscript_info/bee_phylo.RData")
d1 = readRDS("../Data/Manuscript_info/bee_data.RData")
n = readRDS("../Data/Manuscript_info/node_data.RData")

#Circular layour
ggtree(tree_family_interactions, size=0.6, open.angle=1, alpha=1)+ 
geom_tippoint(aes(size= Interactions, fill = Interactions), color="black", shape=21) +
geom_tiplab(linetype='dashed', linesize=.05, offset = 0.063, size=2.9, fontface=1) +
scale_size(range = c(0,4.5), guide = 'none')  +
scale_fill_continuous(type = "viridis", direction=-1) + 
guides(fill=guide_legend(), size = guide_legend()) +
theme(legend.position = "bottom") +
new_scale_fill() +
geom_fruit(data = d1, geom=geom_bar,
                    mapping=aes(y=Genus1, x=Percent, fill = fam_group),
                    pwidth=0.13, 
                    orientation="y", 
                    stat="identity",
                    color="black",
                    offset = 0.137,
            axis.params=list(
                         axis = "x",
                         text.size  = 2.5,
                         hjust  = 1,
                         vjust = 0.5,
                         nbreak = 3,
                         fontface = 2
                     ),
         grid.params=list()) +
guides(fill = "none")+
scale_fill_manual(values = c(a = "tan2", b = "lightgrey"), guide = "none")+
geom_fruit(geom=geom_bar,
                    mapping=aes(y=label, x=Spp_number_Europe),
                    pwidth=0.12, 
                    orientation="y", 
                    stat="identity",
                    color="black",
                    offset = 0,
            axis.params=list(
                         axis = "x",
                         text.size  = 2.5,
                         hjust  = 1,
                         vjust = 0.5,
                         nbreak = 1,
                         fontface = 2
                     ),
         grid.params=list()) +
annotate("text", x = 14.2, y = 70, label = "SafeNet spp\n coverage (%)", size= 3, fontface= "bold") +
annotate("text", x = 15.7, y = 70, label = "European\n species", size= 3, fontface= "bold")+
annotate("text", x = 12.1, y = 70, label = "SafeNet\n interactions", size= 3, fontface= "bold")+ 
new_scale_fill() +
geom_hilight(data=n, aes(node=node), type = "roundrect", fill = "orange") +
annotate("text", x = 7.5, y = 2.8, label = "Melittidae", size= 3)+
annotate("text", x = 7, y = 6.5, label = "Andrenidae", size= 3)+
annotate("text", x = 9.5, y = 12.5, label = "Colletidae", size= 3)+
annotate("text", x = 5.9, y = 17.4, label = "Halictidae", size= 3)+
annotate("text", x = 2.8, y = 28.6, label = "Apidae", size= 3)+
annotate("text", x = 2.8, y = 51.3, label = "Megachilidae", size= 3) +
coord_cartesian(clip = "off")+
theme(legend.margin=unit(1,"cm"),
      legend.spacing = unit(1, "cm"),
      plot.margin = margin(4,0,4,0), 
      legend.position = c(0.6,0),
      legend.direction = "horizontal")

```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 3}. Phylogenetic representation of the plant families in the SafeNet database within the context of all flowering plant families occurring in Europe. The number of interactions recorded per genus in the database is illustrated using circles, with their sizes proportional to the number of interactions on a logarithmic scale. Additionally, a gradient of colors ranging from yellow to dark purple aids in this visualization. The coverage of species recorded in SafeNet per genus at the European level is depicted with orange and light grey bars, representing the percentage of species included and not included in the database, respectively, out of the total number of flowering plant species per family at the European level. Grey bars indicate the total number of species per family at the European level on logarithmic scale.", fig.width=8, fig.height=8}

@@

#Save data and plot it in an Rmarkdown
tree_family_interactions = readRDS("../Data/Manuscript_info/plant_phylo.RData")
d1 = readRDS("../Data/Manuscript_info/plant_coverage_phylo_RData")

#Check if there are differences in labels and order
ggtree(tree_family_interactions, layout="circular", size=0.15, open.angle=5, alpha=0.25) +
geom_tippoint(aes(size= Interactions, fill = Interactions), color="black", shape=21) +
geom_tiplab(linetype='dashed', linesize=.05, offset = -5, size=1.85, fontface=2, hjust=1) +
scale_size(range = c(0,4), guide = 'none') +
scale_fill_continuous(type = "viridis", direction=-1) + 
theme(legend.position = "bottom") +
guides(fill=guide_legend(), size = guide_legend()) +
new_scale_fill() +
geom_fruit(data = d1, geom=geom_bar,
   mapping=aes(y=label, x=Percent, fill = fam_group),
   pwidth=0.1, 
   orientation="y", 
   stat="identity",
   color="black",
   offset = 0.028,
   linewidth=0.35,
   axis.params=list(
     axis = "x",
     text.size  = 1.5,
     hjust  = 1,
     vjust = 0.5,
     nbreak = 3,
     fontface = 2),
  grid.params=list())+
guides(fill = "none")+
scale_fill_manual(values = c(A = "lightgrey", B = "tan2"), guide = "none") +
geom_fruit(geom=geom_bar,
    mapping=aes(y=label, x=log(Spp_Europe +1)),
    pwidth=0.1, 
    orientation="y", 
    stat="identity",
    color="black",
    offset = 0,
    linewidth=0.35) +
theme(legend.margin=unit(1,"cm"),
      legend.spacing = unit(1, "cm"),
      plot.margin = margin(4,0,4,0), 
      legend.position = c(0.5,0.09),
      legend.direction = "horizontal")
